---
title: "SI 699 Demo"
author: "Team 1"
date: "2024-03-012"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro to R 

```{r}
# install.packages('tidyverse')
library(tidyverse)
```


```{r}
data('iris')
head(iris)
```

## Running A Regression 

Logistic Regression
```{r}
iris <- iris %>% mutate(is_setosa = ifelse(Species == 'setosa', 1, 0))
```

```{r}
logistic_regression_model <- glm(is_setosa ~ Sepal.Length, data = iris, family = 'binomial')
```

```{r}
summary(logistic_regression_model)
```
## Multinomial Logistic Regression (Discrete Choice Model)

*Logistic regression*: Models binary outcomes.
*Multinomial logistic regression*: The model to use when your outcome variable takes one of several options.
- **Example**: A consumer has the option to purchase a product from one of several brands.

### Applications
- Market share and pricing analysis (how would increasing prices affect market share?)
- New technology adoption (what factors influence the adoption of a new technology?)
- Competitive analysis (how do product/service features influence consumer preferences?)

```{r}
library(mlogit) # Note the multinom function from the nnet package does not require reshaping dataMi
library(caret)
print('hello world')

```


