---
title: "MRP Prototype"
author: "Haley Johnson"
date: "2024-02-20"
output: html_document
---

## Run with DF3, m4, post strat state and then the last two things for pred. I still might re run with more data. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r}
library(arrow)
library(broom.mixed)
library(gutenbergr)
library(haven)
library(knitr)

library(labelled)
library(modelsummary)
library(rstanarm)
library(ggalt)
library(bayesplot)
library(tidybayes)
library(tidyverse)
library(usmap)
library(scales)
```

## Load Data

### Monmouth Polls
```{r}
monmouth <- read.csv("../data/nat_2020_mar_may_june_august.csv")
```

```{r}
monmouth <- monmouth %>% rename(STATEFIP = state)
```

```{r}
monmouth <- monmouth %>% mutate(age_recoded = factor(age_recoded, levels = c(1, 2, 3)), 
                    education_recoded = factor(education_recoded, levels = c(1, 3)), 
                    race_recoded = factor(race_recoded, levels = c(1, 2, 3, 4, 9)), 
                    #male = as.factor(male),
                    STATEFIP = as.factor(STATEFIP))
```

```{r}
monmouth <- monmouth %>% filter(vote_choice_recoded == 1 | vote_choice_recoded == 0)
monmouth <- monmouth %>% mutate(vote_choice_recoded = as.factor(vote_choice_recoded))
```

### Comet Polls
```{r}
comet <- read.csv("../data/comet_recodedv2.csv")
```

```{r}
comet <- comet %>% mutate(
  male = as.factor(gender_coded), 
  education_recoded = factor(education_coded, levels = c(1, 3)), 
  age_recoded = factor(age_group, levels = c(1, 2, 3)), 
  race_recoded = factor(race_coded, levels = c(1, 2, 3, 4, 9)), 
  STATEFIP = as.factor(STATEFP), 
  vote_choice_recoded = as.factor(vote_coded)) %>%
  select(male, education_recoded, age_recoded, race_recoded, STATEFIP, vote_choice_recoded)
```

### Harvard Election Survey 
```{r}
harvard <- read.csv('../data/harvard_poll.csv')
```

```{r}
harvard <- harvard %>% 
  mutate(age_recoded = as.factor(age_recoded), 
         #male = as.factor(male), 
         education_recoded = as.factor(education_recoded), 
         race_recoded = as.factor(race_recoded), 
         vote_choice_recoded = as.factor(vote_choice_recoded), 
         STATEFIP = as.factor(STATEFIP))
```


### State Info
```{r}
extra_info <- read.csv("../data/nat_2020_june_special.csv")
```

```{r}
extra_info <- extra_info %>% 
              select(rating, STATEFIP, e_votes, state, dem, republican, dem.advantage) %>%
              unique() %>%
              mutate(STATEFIP = as.factor(STATEFIP), 
                     rating = as.factor(rating))
```


### Post Stratification Data
```{r}
post_strat <- read.csv("../data/census/cleaned/post_stratification_data_by_state.csv")

post_strat <- post_strat %>% mutate(age_recoded = as.factor(age_recoded), 
                      race_recoded = as.factor(race_recoded), 
                      STATEFIP = as.factor(STATEFIP),
                      education_recoded = as.factor(education_recoded))
```

```{r}
post_strat$male[which(post_strat$male == 'True')] = 1
post_strat$male[which(post_strat$male == 'False')] = 0

post_strat <- post_strat %>% mutate(male = factor(male, levels = c(1, 0)))
```

```{r}
post_strat <- left_join(post_strat, extra_info, by = 'STATEFIP')
```

```{r}
tot = sum(post_strat$PERWT)
post_strat <- post_strat %>% mutate(pop_prop = PERWT / tot)
```

### New Data Source
```{r}
new = read.csv("../data/nat_2024_to_pred.csv")
new <- new %>% 
        select(age_recoded, education_recoded, race_recoded, male, STATEFIP, vote_choice_recoded) %>%
        mutate(age_recoded = as.factor(age_recoded), 
               education_recoded = as.factor(education_recoded), 
               race_recoded = as.factor(race_recoded), 
               male = as.factor(male),
               vote_choice_recoded = as.factor(vote_choice_recoded),
               STATEFIP = as.factor(STATEFIP))
```

```{r}
df = left_join(new, extra_info, by = 'STATEFIP')
```


### Combine Data Sources
```{r}
df <- rbind(monmouth, harvard)
df <- left_join(df, extra_info, by = 'STATEFIP')
```

```{r}
df$male[which(df$male == 'True')] = 1
df$male[which(df$male == 'False')] = 0
df <- df %>% mutate(male = factor(male))
```

### Reweight Survey
Add weights to be representative of the voting age population 

```{r}
big_n = nrow(df)
sample_props <- df %>% 
                  group_by(education_recoded, age_recoded, race_recoded, male) %>%
                  summarise(sample_prop = n() / big_n)

pop_props <- post_strat %>% 
              group_by(education_recoded, age_recoded, race_recoded, male) %>%
              summarise(pop_prop = sum(PERWT) / tot)
  
```

```{r}
reweight <- left_join(sample_props, pop_props, by = c('age_recoded', 'race_recoded', 'male', 'education_recoded'))
```

Caps weights between 0.5 and 2 to keep estimates numerically stable. 
```{r}
weights <- reweight %>%
              select(age_recoded, race_recoded, education_recoded, male, sample_prop, pop_prop) %>%
              unique() %>%
              mutate(weight = sample_prop/pop_prop) %>%
              mutate(weight = ifelse(weight > 2, 2, weight)) %>%
              mutate(weight = ifelse(weight < 0.5, 0.5, weight)) %>%
              select(age_recoded, race_recoded, education_recoded, male, weight)
```

```{r}
df <- left_join(df, weights, by = c('age_recoded', 'race_recoded', 'male', 'education_recoded'))
post_strat <- left_join(post_strat, weights, by = c('age_recoded', 'race_recoded', 'male', 'education_recoded'))
```


## MRP
```{r}
m1 <- stan_glmer(vote_choice_recoded ~ (1|male) + (1|STATEFIP) + (1 |education_recoded) + (1 |age_recoded)  + (1|race_recoded) + rating + dem.advantage + republican + weight,
                 data = df , 
                 family = binomial(link = "logit"), 
                 prior = normal(location = 0, scale = 0.5, autoscale = TRUE), 
                 prior_covariance = decov(scale = 0.50),
                 adapt_delta = 0.99,
                 prior_intercept = normal(location = 0, scale = 0.5, autoscale = TRUE), 
                 seed = 13)
 

```

##Get Propensity Scores
```{r}
epred_mat <- posterior_epred(m1, newdata = post_strat, draws = 2500)
mrp_estimates_vector <- epred_mat %*% post_strat$PERWT / sum(post_strat$PERWT)
mrp_estimate <- c(mean = mean(mrp_estimates_vector), sd = sd(mrp_estimates_vector))
cat("MRP estimate mean, sd: ", round(mrp_estimate, 3))
```



```{r}
subgroups_df <- df %>% 
  select(age_recoded, race_recoded, male, education_recoded, STATEFIP)  %>% 
  expand(age_recoded, race_recoded, male, education_recoded, STATEFIP)  %>%
  mutate(male = factor(male, levels = c(1,0))) %>%
  mutate(mrp_subgroup_estimate = NA, mrp_subgroup_estimate_se = NA)


# Loop to populate the dataframe
for(i in 1:nrow(subgroups_df)) {
  # Filtering and poststratification
  filtering_condition <- which(post_strat$male == subgroups_df$male[i] & 
                                 post_strat$age_recoded == subgroups_df$age_recoded[i] & 
                                 post_strat$race_recoded == subgroups_df$race_recoded[i] &    
                                 post_strat$education_recoded == subgroups_df$education_recoded[i] & 
                                 post_strat$STATEFIP == subgroups_df$STATEFIP[i])

  epred_mat_filtered <- epred_mat[, filtering_condition]
  k_filtered <- post_strat[filtering_condition, ]$PERWT
  mrp_subgroup_estimates_vector <- (epred_mat_filtered * k_filtered) / sum(k_filtered)
  subgroups_df$mrp_subgroup_estimate[i] <- mean(mrp_subgroup_estimates_vector)
  subgroups_df$mrp_subgroup_estimate_se[i] <- sd(mrp_subgroup_estimates_vector)
}
```


```{r}
write_csv(subgroups_df, "new_prop_scores_normalize_it_FINAL.csv")
```


## Visualizations

### Coefficents
```{r}
m1 %>% 
  tidy() %>%
  ggplot() + 
  geom_bar(aes(x = estimate, y = term), stat = 'identity') + 
  labs(x = 'Feature', y = 'Coefficient', title = 'Regression Coefficent')
```


### Overall Conditional Means
```{r}
m1 |>
  spread_draws(`(Intercept)`, b[, group]) |>
  mutate(condition_mean = `(Intercept)` + b) |>
  separate(col = group, 
           into = c("type", "instance"), 
           sep = ":", remove = FALSE) |> 
  filter(type != "STATEFIP") |> 
  ggplot(aes(y = group, x = condition_mean)) +
  stat_halfeye() +
  scale_y_discrete(labels = c('18-34', '33-55', '56+', 'Beyond HS Education', 
                              'Less than HS Education', 'Female', 'Male', 
                              'White', 'Black', 'Asian', 'Latino', 'Race Other')) +
  labs(x = 'Conditional Mean', y = '', title = 'MRP Feature Importance') + 
  theme_bw()

ggsave("beyonce_hs_education.png", bg = "transparent")
```

### Conditional Means for Sex
```{r}
m1 |>
  spread_draws(`(Intercept)`, b[, group]) |>
  mutate(condition_mean = `(Intercept)` + b) |>
  separate(col = group, 
           into = c("type", "instance"), 
           sep = ":", remove = FALSE) |> 
  filter(type == "male") |> 
  ggplot(aes(y = group, x = condition_mean)) +
  stat_halfeye() +
  theme_minimal()
```

### Conditional Means for Race

```{r}
m1 |>
  spread_draws(`(Intercept)`, b[, group]) |>
  mutate(condition_mean = `(Intercept)` + b) |>
  separate(col = group, 
           into = c("type", "instance"), 
           sep = ":", remove = FALSE) |> 
  filter(type == "male") |> 
  ggplot(aes(y = group, x = condition_mean)) +
  stat_halfeye() +
  theme_minimal()
```

### Survey Representatitiveness DF
```{r}
sample_overall_n = nrow(df)
pop_overall_n = sum(post_strat$PERWT)
```


#### Racial Representativeness 
```{r}
race_props <- post_strat %>% 
  group_by(race_recoded) %>% 
  summarise(subset_pop_prop = sum(PERWT) / pop_overall_n)
```

```{r}
df %>% 
  group_by(race_recoded) %>%
  summarise(n = n(), sample_prop = n() / sample_overall_n) %>%
  left_join(y = race_props, by = c('race_recoded')) %>%
  select(sample_prop, race_recoded, subset_pop_prop) %>% 
  ggplot(aes(y = factor(race_recoded, levels = c(9, 3, 2, 4, 1)), x = sample_prop, xend = subset_pop_prop)) +
  geom_dumbbell(size = 4, color="#e3e2e1", colour_x = "#3e45c9", colour_xend = "#e32245", dot_guide = TRUE, dot_guide_size = 0.25) + 
  ggtitle('Racial Composition of Polls vs National Voting Age Population') + 
  xlab('Proportion') + 
  ylab('Racial or Ethnic Group') + 
  scale_x_continuous(labels = scales::percent) + 
  scale_y_discrete(labels = c('1' = 'White', '2' = 'Black', '3' = 'Asian', '4' = 'Hispanic or Latino', '9' = 'Other')) + 
  theme_minimal() + 
  theme(panel.grid.major.x = element_line(linewidth = 0.05))
```

#### State Representativeness

```{r}
state_props <- post_strat %>% 
  group_by(STATEFIP, state) %>% 
  summarise(subset_pop_prop = sum(PERWT) / pop_overall_n)

df %>% 
  mutate(state = tolower(state)) %>%
  group_by(STATEFIP) %>%
  filter(STATEFIP %in% c(4, 6, 12,13, 26, 32, 36, 42, 48, 51, 55)) %>%
  summarise(n = n(), sample_prop = n() / sample_overall_n) %>%
  left_join(y = state_props, by = c('STATEFIP')) %>%
  select(sample_prop, state, subset_pop_prop) %>% 
  ggplot(aes(y = state, x = sample_prop, xend = subset_pop_prop)) +
  geom_dumbbell(size = 2, color="#e3e2e1", colour_x = "#3e45c9", colour_xend = "#e32245", dot_guide = TRUE, dot_guide_size = 0.15) + 
  scale_x_continuous(labels = scales::percent) + 
  scale_y_discrete(limits = rev) + 
  ggtitle('State Composition of Poll vs National Voting Age Population') + 
  xlab('Proportion') + 
  ylab('State') + 
  theme_minimal() + 
  theme(panel.grid.major.x = element_line(linewidth = 0.05), axis.text = element_text(size = 7))

ggsave("test_2.png", bg = "transparent")

png(filename="test.png", height = 4000, width = 6000, res=400) 
par(bg=NA)
plot(state_vs_real)
dev.off()
```


#### Age Representativeness
```{r}
age_props <- post_strat %>% 
  group_by(age_recoded) %>% 
  summarise(subset_pop_prop = sum(PERWT) / pop_overall_n)


#png(filename="bench_query_sort.png", width=600, height=600)
df %>% 
  group_by(age_recoded) %>%
  summarise(n = n(), sample_prop = n() / sample_overall_n) %>%
  left_join(y = age_props, by = c('age_recoded')) %>%
  select(sample_prop, age_recoded, subset_pop_prop) %>% 
  ggplot(aes(y = age_recoded, x = sample_prop, xend = subset_pop_prop)) +
  geom_dumbbell(size = 3, color="#e3e2e1", colour_x = "#3e45c9", colour_xend = "#e32245", dot_guide = TRUE, dot_guide_size = 0.15) + 
  scale_x_continuous(labels = scales::percent) + 
  scale_y_discrete(limits = rev, labels = c('1' = '18-34', '2' = '34-55', '3' = '55+')) + 
  ggtitle('Age Composition of Poll vs National Voting Age Population') + 
  xlab('Proportion') + 
  ylab('Age') + 
  theme_minimal() + 
  theme(panel.grid.major.x = element_line(linewidth = 0.05), axis.text = element_text(size = 7))
```

### Effect of Propensity Estimates

```{r}
state_subgroups_df <- df %>% select(STATEFIP)  %>% 
  left_join(extra_info[, c('STATEFIP', 'state')], by = 'STATEFIP') %>%
  unique() %>%
  mutate(mrp_subgroup_estimate = NA, mrp_subgroup_estimate_se = NA)

for(i in 1:nrow(state_subgroups_df)) {
  filtering_condition <- which(post_strat$STATEFIP == state_subgroups_df$STATEFIP[i])
  epred_mat_filtered <- epred_mat[, filtering_condition]
  k_filtered <- post_strat[filtering_condition, ]$PERWT
  mrp_subgroup_estimates_vector <- (epred_mat_filtered %*% k_filtered) / sum(k_filtered)
  state_subgroups_df$estimate[i] <- mean(mrp_subgroup_estimates_vector)
  state_subgroups_df$estimate_se[i] <- sd(mrp_subgroup_estimates_vector)

}
```

```{r}
state_support <- df %>% 
  group_by(state) %>%
  summarise(estimate = mean(as.integer(vote_choice_recoded) - 1, na.rm = TRUE), 
            estimate_sd = sd(as.integer(vote_choice_recoded) - 1), 
            n = n(),
            estimate_se = estimate_sd / sqrt(n)) 
```

```{r}
state_support$source = 'polls'
state_subgroups_df$source = 'mrp'
ests <- rbind(state_support[, c('state', 'estimate', 'estimate_se', 'source')], 
              state_subgroups_df[, c('state', 'estimate', 'estimate_se', 'source')])
```


```{r}
ests %>% 
  #filter(state %in% c('michigan', 'georgia', 'wisconsin', 'pennsylvania', 'arizona', 'texas', 'california'))) %>%
  ggplot() + 
  geom_point(aes(x = state, y = estimate, color = source)) + 
  geom_errorbar(aes(ymin = estimate - 2 * estimate_se, ymax = estimate + 2 * estimate_se, x = state, color = source), 
                alpha = 0.5, width = 0) + 
  scale_y_continuous(breaks=c(0,.25,.5,.75,1), 
                     labels=c("0%","25%","50%","75%","100%"), 
                     expand=c(0,0))+
  coord_cartesian(ylim=c(0, 1)) +
  theme_bw() +
  labs(x="States",y="Support", title='MRP Estimates vs Poll Averages by State') +
  theme(axis.title=element_text(size=10),
        axis.text.y=element_text(size=10),
        axis.text.x=element_text(angle=90,size=7, vjust=0.3),
        legend.title=element_text(size=10),
        legend.text=element_text(size=10))
ggsave("test_3.png", bg = "transparent")
```


